---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>"
)
library(iterake)
```

# iterake <img src=logo/ITERAKE_LOGO_01.png width=140 height=140 align="right" />  
**NOTE**: This package is under active development, use at your own risk.
  
## Overview 

iterake's main utility is creating row-level weights using a process called iterative raking. Iterative raking (also known as rim weighting), is one of several methods used to correct the deviation between the *marginal* proportions in a sample and a known population, or, universe as it was first referred to (Deming & Stephan 1940) for a given set of variables.

The weighting process with `iterake` is fairly straightforward; we suggest the following workflow: 

1. Use the `universe()` function to build your population. 
    1. The univerise is constructed with one or more categories where the marginal probabilites are known. These categories are built with the `category()` function.
2. Compare the marginal proportions in your sample with the population with `compare_margins()` function.  
3. If needed, create weights for your data using `iterake()`.  
4. Verify that the weighted proportions in your sample now match the population. The `compare_margins()` is used here again.  
5. Check the performance of the weighting model with `weight_stats()`.

## Installation
``` {r, eval = FALSE}
# Install the development version from GitHub
install.packages("devtools")
devtools::install_github("ttrodrigz/iterake")
```


## Motivating Example
Suppose a car dealership is interested in surveying the opinions of people who purchased a new vehicle from their lot. This dealership has been around since 2015 and has kept accurate records of its sales since its inception. The dealership randomly samples 400 individuals from its sales database to complete a customer satisfaction survey.  

The dealership wants to make sure their sample is representative of the population, so they
calculate the marginal proportions of a few key categories:

```{r, echo=FALSE, results='asis'}
age <- data.frame(
    Age    = c("18-34", "35-54", "55+"),
    Pop.   = c("12%", "58%", "30%"),
    Sample = c("16%", "55.5%", "28.5%")
)
year <- data.frame(
    Year   = c(2015, 2016, 2017, 2018),
    Pop.   = c("22%", "25%", "32%", "21%"),
    Sample = c("19%", "21%", "36%", "24%")
)
vehicle <- data.frame(
    Type   = c("Car", "SUV", "Truck"),
    Pop.   = c("38%", "47%", "15%"),
    Sample = c("36%", "52%", "12%")
)

knitr::kable(list(age, year, vehicle), format = "html", align = "c")
```

There are slight deviations between what was collected in the sample and the known population. These deviations are small enough where weighting can be performed to correct for these differences between the marginal proportions.

#### **Step 1**: Build the universe with known marginal proportions based on categories of interest.
```{r, include=TRUE, cache=TRUE, cache.lazy=TRUE, eval=TRUE}
# load iterake and the data
library(iterake)
data("dealer_data")

# build the 'universe'
dealer_uni <- universe(
    
    df = dealer_data,
    
    category(
        name = "Age",
        buckets = c("18-34", "35-54", "55+"),
        targets = c(.12, .58, .30)
    ),
    
    category(
        name = "Year",
        buckets = c(2015, 2016, 2017, 2018),
        targets = c(.22, .25, .32, .21)
    ),
    
    category(
        name = "Type",
        buckets = c("Car", "SUV", "Truck"),
        targets = c(.38, .47, .15)
    )
    
)
```  

#### **Step 2**: Compare *unweighted* sample proportions to the universe.  
```{r}
compare_margins(
    df = dealer_data,
    universe = dealer_uni
)
```

#### **Step 3**: Weight the data.
```{r}
dealer_weighted <- iterake(
    df = dealer_data, 
    universe = dealer_uni
)

dealer_weighted
```

#### **Step 4**: Compare *weighted* sample proportions to the universe.  
````{r}
compare_margins(
    df = dealer_weighted,
    universe = dealer_uni,
    weight = weight
)
```

#### **Step 5**: Inspect performance of weighting model.
```{r}
weight_stats(dealer_weighted$weight)
````