---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  comment = "#>"
)
library(iterake)
```

# iterake <img src=logo/ITERAKE_LOGO_01.png width=140 height=140 align="right" />  

## Overview 

iterake's main utility is creating row-level weights using a process called iterative raking. Iterative raking (also known as rim weighting), is one of several methods used to correct the deviation between the *marginal* proportions in a sample and a known population, or, universe as it was first referred to (Deming & Stephan 1940) for a given set of variables.  

iterake is designed with speed and simplicity in mind. The weighting algorithm is powered by [data.table](https://github.com/Rdatatable/data.table/wiki) and takes advantage of its fast [grouping](https://github.com/Rdatatable/data.table/wiki/Benchmarks-:-Grouping) and joining; it tends to scale well even with larger data sets. The functions of this package are designed to play nicely users of the  [tidyverse](https://github.com/tidyverse/tidyverse). Columns in the data can be accessed without quotation, and every function returns a [tibble](https://github.com/tidyverse/tibble).  

## Workflow  

The weighting process with `iterake` is fairly straightforward, we suggest:  

1. Use the `universe()` function to build your population. 
    1. The univerise is constructed with one or more categories where the marginal probabilites are known. These categories are built with the `category()` function.
    2. If you want to use the natural marginal proportions from an existing dataset as your targets, you can use `inherit_category()`. Just make sure the name given to the category matches the existing data and the data you intend to weight.
2. Compare the marginal proportions in your sample with the population with `compare_margins()` function.  
3. If needed, create weights for your data using `iterake()`.  
4. Use `compare_margins()` again to verify that the weighted proportions in your sample now match the population.  
5. Check the performance of the weighting model with `weight_stats()`.  
  
## Installation
``` {r, eval = FALSE}
# Install the development version from GitHub
install.packages("devtools")
devtools::install_github("ttrodrigz/iterake")
```

## Motivating Example
Suppose a car dealership is interested in surveying the opinions of people who purchased a new vehicle from their lot. This dealership has been around since 2015 and has kept accurate records of its sales since its inception. The dealership randomly samples 400 individuals from its sales database to complete a customer satisfaction survey.  

The dealership wants to make sure their sample is representative of the population, so they calculate the marginal proportions of a few key categories:

```{r, echo=FALSE, results='asis'}
age <- data.frame(
    Age    = c("18-34", "35-54", "55+"),
    Pop.   = c("12%", "58%", "30%"),
    Sample = c("16%", "55.5%", "28.5%")
)
year <- data.frame(
    Year   = c(2015, 2016, 2017, 2018),
    Pop.   = c("22%", "25%", "32%", "21%"),
    Sample = c("19%", "21%", "36%", "24%")
)
vehicle <- data.frame(
    Type   = c("Car", "SUV", "Truck"),
    Pop.   = c("38%", "47%", "15%"),
    Sample = c("36%", "52%", "12%")
)

knitr::kable(list(age, year, vehicle), format = "html", align = "c")
```

There are slight deviations between what was collected in the sample and the known population. These deviations are small enough where weighting can be performed to correct for these differences between the marginal proportions.

#### **Step 1**: Build the universe with known marginal proportions based on categories of interest.
```{r, include=TRUE, cache=TRUE, eval=TRUE}
# data comes pre-packaged with iterake
library(iterake)
data("dealer_data")

# build the 'universe'
dealer_uni <- universe(
    
    df = dealer_data,
    
    category(
        name = "Age",
        buckets = c("18-34", "35-54", "55+"),
        targets = c(.12, .58, .30)
    ),
    
    category(
        name = "Year",
        buckets = c(2015, 2016, 2017, 2018),
        targets = c(.22, .25, .32, .21)
    ),
    
    category(
        name = "Type",
        buckets = c("Car", "SUV", "Truck"),
        targets = c(.38, .47, .15)
    )
    
)
```  

#### **Step 2**: Compare *unweighted* sample proportions to the universe.  
```{r}
compare_margins(dealer_data, universe = dealer_uni)
```

#### **Step 3**: Weight the data.
```{r}
dealer_weighted <- iterake(dealer_data, dealer_uni)

dealer_weighted
```

#### **Step 4**: Compare *weighted* sample proportions to the universe.  
````{r, message=FALSE, fig.width = 10, fig.height = 4}
library(dplyr)

# re-inspect proportions with plot
compare_margins(
    df = dealer_weighted, 
    weight = weight, 
    universe = dealer_uni,
    plot = TRUE
) %>%
    select(-contains("uwgt"))
```

#### **Step 5**: Inspect performance of weighting model.
```{r}
# want to make sure weighting model is efficient 
# and there aren't any extreme weights

weight_stats(dealer_weighted$weight)
````

## Final Notes
* A big **THANK YOU** goes out to our friend James Bouchey for designing the `iterake` hex logo. Please check out his [portfolio](https://jamesbouchey.myportfolio.com/) as his talents extend far past making super clever package logos.  
* We are currently preparing to submit iterake to CRAN, so any feedback is well-received.  
* Stay tuned for a [pkgdown](https://github.com/r-lib/pkgdown) site for this package with more details on how, why, and when to use iterake.
